Active deformation of the Shargyn Basin, a transpressional strike-slip intersection in western Mongolia.

### 2013-05-09 writing/paper/abstract.tex
almost done

\begin{abstractpage}
Intraplate faulting in central Asia is a major component of the Indo-Eurasian collision. The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources. We examine the fault kinematics surrounding the 150km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system. The studies were performed using satellite data and targeted field transects. The results suggest the Shargyn basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many strike-slip terminating thrust splays. Furthermore, local foliation is almost always fault parallel, pointing to the importance of pre-existing structural weaknesses in the development of active faults. This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.
\end{abstractpage}

>>>>

\begin{abstractpage}
Intraplate faulting in central Asia is a major component of the Indo-Eurasian collision. The kinematics and mechanisms of intraplate deformation are important in understanding broad active tectonic patterns, reconstructing past tectonics, analyzing seismic hazard and identifying potential resources. We examine the fault kinematics surrounding the 150 km wide Shargyn Basin at the intersection of the left-lateral transpressional Gobi-Altai fault system and the right-lateral transpressional Altai fault system. The studies were performed using satellite data and targeted field transects. The results suggest the Shargyn basin is formed by a compressional stepover, an uplifted wedge from the intersecting strike-slip systems and many strike-slip terminating thrust splays. Furthermore, local foliation is almost always fault parallel, pointing to the importance of pre-existing structural weaknesses in the development of active faults. This research demonstrates some of the potential kinematics for intraplate transpressional orogenies and emphasizes the importance of pre-existing crustal structure in the development of active faults.
\end{abstractpage}

### 2013-05-09 writing/paper/background.tex
almost done

	Most of the active deformation in Mongolia is probably a result of the Indo-Eurasian collision \citep{Yin2010}\citep{Walker2007}\citep{Cunningham2005a}\citep{England1997}. Indo-Eurasian collisional deformation processes can be separated into "kinematic" and "dynamic" models. The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citet{Tapponnier1982}\citep{Peltzer1996}. This hypthesis suggests that most motion has been accomodated by large strike-slip and thrust faults. The "dynamic" models propose that the viscosity of the lithosphere is more important than upper crustal faults \citep{Houseman1993}\citep{England1997}. In these models, the important effects continuous viscous deformation of the lithosphere and the gravitational effects of the resulting thickening.

	Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}. The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponnier1982}\citep{Peltzer1996}\citep{Houseman1993}. But, because of its orientation, active EASS rollback would exert stresses in a similar orientation to the Indo-Eurasian collision. However, the potential for far-field stresses from subduction zone rollback is not well constrained. Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins. Much of the EASS was experiencing rollback during the late Cretaceous until the Mid-Miocene \citep{Northrup1995}\citep{Schellart2005a}. A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000km from the subduction zone. Additional constraints on the active and past deformation in western Mongolia could influence these broad theories of Asian deformation.

	 The intensity of intracontinental deformation in Mongolia is emphasized by measured GPS motions and three of the largest strike-slip earthquakes ever. For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high. GPS measurements (Figure \ref{broad}) show that ongoing motion is to the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}. The average rate of motion is approximately 5 mm/year relative to a stable Eurasia as defined by \citet{Calais2003}. Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century. In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia. The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}. The Bogd Earthquake has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.  The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.

@@

	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}. The region is dominated by NW-SE striking right-lateral slip. The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults. Others are linked by a right-slip faults. In some places, the thrusts are part of a restraining bend. Major faults, GPS motions, and earthquake moment tensors are shown in Figure \ref{broad}. Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}. The geologic slip rate for a major transpressional fault, the Hoh-Serh-Tsagaan-Salaa Fault, is ~0.3 mm/yr of shortening and ~0.9 mm/yr of dextral slip. This consitutes ~20\% of the total strain accumulation in the Altai. \citep{Frankel2010}

	Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}. North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures. The deformation is diffuse over a 250-350km north-south extent. The range accomodate the eastwards and northeastwards deformation as measured by GPS. At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan Some estimates suggest 30-40km of Cenozoic slip on the Gobi-Tien Shan Fault System (shown as "GBTS" in Figure \ref{asia}) \citep{Cunningham2003a}. There are no slip estimates for the Bogd Fault System in the north. However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10km of slip \citep{Cowie1992}.  Earthquake moment tensors show mainly left-lateral strike-slip in the Gobi-Altai.

	Previous researchers have suggested that the Hangay Dome in central Mongolia (Figure \ref{broad}) is a rigid block causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems) \citep{Cunningham2005a}. However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome. This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south, and may not be as rigid as previously thought\citep{Walker2006}\citep{Walker2007}\citep{Walker2008}.

@@

	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system. It is approximately 150km in east-west extent and 100km in north-south extent. A small ephemeral lake is located in the center of the basin. The basin margin varies between 1500m and 2000m, while the center is at an elevation of 963m. The basin is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}. To the east and south, the tectonic regime is less certain. The Khantaishir Range to east has been the subject of petrology studies concerning the ophiolite located there. However, the mapping from these studies has not seriously examined the potentially active faults bounding the range to the southwest and west towards the Shargyn Basin.  The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, are also poorly understood.

@@

Many different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}. Some of the settings we might expect:
\begin{itemize}
	\item Foreland basins form in front of a simple thrust system. Flexural subsidence can result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure. Shown in Figure \ref{basintypes}C
	\item Compresional stepover basins form similarly to restraining bend basins. However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault. In the compressional case, the basin is essentially a foreland basin in front of the thrust. Arguably, these fault systems are just large restraining bends. Shown in Figure \ref{basintypes}D.
	\item An intersection wedge can form at a conjugate strike-slip intersection. Because strike-slip motion on both faults must be transferred into uplift, an uplifted wedge of material forms between the two faults. 
\end{itemize}

@@

	To the east of the Sutai Range, the Dariv Thrust uplifts the Dariv Mountains. \citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust. The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range. Southeast of the Dariv Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{grain.png}
  \caption{Schematic map of the dominant structural grain in western Mongolia. The study area is boxed for reference. The shaded Junggar Block is mechanically rigid. The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.  The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Centrail Asian Orogeny. It is not shaded because it is probably not so mechanically rigid.  Modified after \citet{Cunningham2005a}}
  \label{grain}
\end{figure}

@@

\clearpage

>>>>

	Most of the active deformation in Mongolia is probably a result of the Indo-Eurasian collision \citep{Yin2010}\citep{Walker2007}\citep{Cunningham2005a}\citep{England1997}. Indo-Eurasian collisional deformation processes can be separated into "kinematic" and "dynamic" models. The basic "kinematic" model suggests that the Indo-Eurasian collision has caused eastwards extrusion of crustal blocks towards the "free" eastern boundary (the East Asian Subduction Zone)\citep{Tapponnier1982}\citep{Peltzer1996}. This hypothesis suggests that most motion has been accomodated by large strike-slip and thrust faults. The "dynamic" models propose that the viscosity of the lithosphere is more important than upper crustal faults \citep{Houseman1993}\citep{England1997}. In these models, the important effects are continuous viscous deformation of the lithosphere and the gravitational effects of the resulting thickening.

	Rollback on the East Asian Subduction System (EASS) provides an additional explanation for active and past motions in Mongolia \citep{Schellart2005}. The EASS has been treated as a free-boundary in many of the East Asian tectonic models\citep{Tapponnier1982}\citep{Peltzer1996}\citep{Houseman1993}. But, because of its orientation, active EASS rollback would exert similarly oriented stresses to the Indo-Eurasian collision. However, the potential for far-field stresses from subduction zone rollback is not well constrained. Back-arc extension may not be limited to the deep, ocean-floor producing, back-arc basins. Much of the EASS was experiencing rollback during the late Cretaceous until the Mid-Miocene \citep{Northrup1995}\citep{Schellart2005a}. A physical model by \citet{Schellart2005} showed that, if the entire East Asian Subduction zone were to rollback, intraplate deformation could occur as far as 3000 km from the subduction zone. Additional constraints on the active and past deformation in western Mongolia could influence these broad theories of Asian deformation.

	 The intensity of intracontinental deformation in Mongolia is emphasized by measured GPS motions and three of the largest strike-slip earthquakes ever. For a region so far from a plate margin, the rate of crustal deformation in Mongolia is very high. GPS measurements (Figure \ref{broad}) show that ongoing motion is toward the north in the Altai, and mostly eastward in the Gobi-Altai\citep{Calais2003}. The average rate of motion is approximately 5 mm/year relative to a stable Eurasia as defined by \citet{Calais2003}. Multiple earthquakes with a moment magnitude of 8 or greater have occured in the last century. In July 1905, earthquakes with mangitude 7.9 and 8.4 occured on the Bolnay Fault in northern Mongolia. The 1957, magnitude ~8.0 earthquake on the Bogd Fault is one of the best exposed strike-slip surface ruptures in the world\citep{Kurushin1998} \citep{Okal1976a} \citep{Pollitz2003}. The Bogd Earthquake has been used as an analogue for potential Southern California earthquake events \citep{Bayarsayhan1996}.  The intensity of deformation makes Mongolia an ideal place to study the mechanics of intracontinental fault systems.

@@

	Recent studies have examined the neotectonics of the Altai range via satellite imagery and targeted field excursions\citep{Cunningham2005a}. The region is dominated by NW-SE striking right-lateral slip. The primary features are strike-slip faults, but much of the topography is created by thrusts oriented 30 degrees off the main strike slip faults. Others are linked by a right-slip faults. In some places, the thrusts are part of a restraining bend. Major faults, GPS motions, and earthquake moment tensors are shown in Figure \ref{broad}. Pure strike-slip, oblique slip and thrust earthquake moment tensors are observed, which indicates that active motion is variable depending on the specific fault \citep{Bayasgalan2005a}. The geologic slip rate for a major transpressional fault, the Hoh-Serh-Tsagaan-Salaa Fault, is ~0.3 mm/yr of shortening and ~0.9 mm/yr of dextral slip. This constitutes ~20\% of the total strain accumulation in the Altai. \citep{Frankel2010}

	Southeast of the Altai, deformation in the Gobi-Altai range is predominantly E-W striking left-lateral slip \citep{Cunningham2010}. North and south dipping thrusts striking NW-SE follow the basement structures, whereas the strike-slip faults cut across the basement structures. The deformation is diffuse over a 250-350 km north-south extent. The range accomodate the eastwards and northeastwards deformation as measured by GPS. At the southern extent of the range, the Gobi-Tian Shan Fault is continuous from the easternmost Tian Shan through the eastern Gobi-Altai, thus supporting the connection between the Gobi-Altai and the Tian Shan. Some estimates suggest 30-40 km of Cenozoic slip on the Gobi-Tien Shan Fault System (shown as "GBTS" in Figure \ref{asia}) \citep{Cunningham2003a}. There are no slip estimates for the Bogd Fault System in the north. However, empirical relationships between fault slip and fault length suggest that the Bogd Fault System has experienced at least 10 km of slip \citep{Cowie1992}.  Earthquake moment tensors show mainly left-lateral strike-slip in the Gobi-Altai.

	Previous researchers have suggested that the Hangay Dome in central Mongolia (Figure \ref{broad}) is a rigid block causing the formation of a pair of conjugate strike-slip faults (Altai and Gobi-Altai fault systems) \citep{Cunningham2005a}. However, recent detailed mapping work has shown the presence of large strike-slip structure cutting across the Hangay Dome \citep{Walker2007}. This indicates that the Hangay Dome is being brittly deformed in a similar fashion to the Gobi-Altai to the south, and may not be as rigid as previously thought\citep{Walker2006}\citep{Walker2007}\citep{Walker2008}.

@@

	The Shargyn basin is located at the intersection of the Altai fault system and the Gobi-Altai fault system. It is approximately 150 km in east-west extent and 100km in north-south extent. A small ephemeral lake is located in the center of the basin. The basin margin varies between 1500 meters and 2000 meters, while the center is at an elevation of 963 meters. The basin is bounded on the west by the right-lateral Tonhil Fault and on the north by the left-lateral Shargyn Fault \citep{Cunningham2003}. To the east and south, the tectonic regime is less certain. The Khantaishir Range to east has been the subject of petrological studies concerning the ophiolite located there. However, the mapping from these studies has not seriously examined the potentially active faults bounding the range to the southwest and west towards the Shargyn Basin.  The neotectonics of the Bogd Range, which bounds the Shargyn basin to the south, are also poorly understood.

@@

Many different types of basins form in strike-slip and compressional settings. Multiple types of basins can form and interact at the intersection of fault systems \citep{Busby1995}. Some of the settings we might expect:
\begin{itemize}
	\item Foreland basins form in front of a simple thrust system. Flexural subsidence can result from the neighboring uplift. This is shown in Figure \ref{basintypes}A. 
	\item Piggyback basins form in the hinterland of a thrust fault. The adjacent uplift provides a sediment source. Shown in Figure \ref{basintypes}B
	\item Restraining bend basins form similarly to foreland basins with deposition in the foreland of thrusts, but often form on both sides of a positive flower thrust structure. Shown in Figure \ref{basintypes}C
	\item Compresional stepover basins form similarly to restraining bend basins. However, rather than a bend in the primary strike-slip system, there are two parallel strike-slip fault segments, separated by at least one thrust fault. In the compressional case, the basin is essentially a foreland basin in front of the thrust. Arguably, these fault systems are just large restraining bends. Shown in Figure \ref{basintypes}D.
	\item An intersection wedge can form at a conjugate strike-slip intersection. Because strike-slip motion on both faults must be transferred into uplift, an uplifted wedge of material forms between the two faults. Shown in Figure \ref{basintypes}E
\end{itemize}

@@

	To the east of the Sutai Range, the Dariv Thrust uplifts the Dariv Mountains. \citet{Howard2006} describe the Dariv Basin as a piggyback basin, riding on top of the block thrust up by the Dariv Thrust. The Dariv Basin could also be described as a restraining bend basin from deposition and subsidence due to the adjacent Sutai Range. On the northern boundary of the Shargyn Basin, the Shargyn fault, a left-lateral strike slip fault, bounds the southern edge of the Dariv Mountains, separating the Shargyn Basin from the Dariv Basin.

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{grain.png}
  \caption{Schematic map of the dominant structural grain in western Mongolia. The study area is boxed for reference. The shaded Junggar Block is mechanically rigid. The Junggar Block may be a trapped block of oceanic crust \citep{Carroll1990a}.  The Hangay Block consists of microcontinental fragments that accreted during the early stages of the Central Asian Orogeny. It is not shaded because it is probably not mechanically rigid.  Modified after \citet{Cunningham2005a}}
  \label{grain}
\end{figure}
\clearpage

### 2013-05-09 writing/paper/discussion.tex
almost done

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
  \caption{Schematic fault map of the Shargyn Basin region. The extent is approximately the same as the extent in Figure \ref{regional}. Description of the structure is in the text. The shaded areas are basins. Unshaded areas are mountainous. The dotted lines represent the dominant foliation strike. }
  \label{schematic}
\end{figure}

@@

	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes parallel or close to parallel to the important fault in the region. The foliations fit well with a model where faults form along structural weaknesses. The foliations could also be explained by dragging along the fault. With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike. Dragging can be clearly in the satellite imagery of the beds along the Bogd Fault in the intersection wedge. However, in other locations, foliations are still fault-parallel 5-10km away from the surface trace of the fault and dragging is not seen. So, dragging may be locally important in the intersection wedge, but, elsewhere, it is more likely that the faults formed along pre-existing structural weaknesses.

>>>>

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 3.5in]{schematic.png}
  \caption{Schematic fault map of the Shargyn Basin region. The extent is approximately the same as the extent in Figure \ref{intro}. Description of the structure is in the text. The shaded areas are basins. Unshaded areas are mountainous. The dotted lines represent the dominant foliation strike. The inset in the upper right is an even simpler description of the interacting stepover and wedge uplift.}
  \label{schematic}
\end{figure}

@@

	In almost all (exluding specific parts of the intersection wedge) of the locations we studied the foliation strikes parallel or close to parallel to the important fault in the region. The foliations fit well with a model where faults form along structural weaknesses. The foliations could also be explained by dragging along the fault. With enough slip, regardless of the original foliation is an area, dragging of the foliation near the fault will lead to approximately fault-parallel strike. Dragging can be clearly seen in the satellite imagery of the beds along the Bogd Fault in the intersection wedge. However, in other locations, foliations are still fault-parallel 5-10km away from the surface trace of the fault and dragging is not seen. So, dragging may be locally important in the intersection wedge, but, elsewhere, it is more likely that the faults formed along pre-existing structural weaknesses.

### 2013-05-09 writing/paper/intro.tex
removed sections on gravity
almost done

	In this paper, I examine the intersection of the Altai Range and the Gobi-Altai Range. The Altai Range and the Gobi-Altai Range are two large, actively forming, mountain ranges in central Asia. The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}. The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history. Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.

	The 150km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system. Here, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics. The kinematics consist of a compressional strike-slip stepover, a conjugate strike-slip intersection and termination thrust splays. I also use gravity data to argue that a difference in basement rock has been a primary control on the neotectonic development of the Shargyn Basin region.

>>>>

	In this paper, I examine the intersection of the Altai Range and the Gobi-Altai Range. The Altai Range and the Gobi-Altai Range are two large, actively forming mountain ranges in central Asia. The Altai Range is characterized by right-lateral transpressional slip, while the Gobi-Altai Range is characterized by left-lateral transpressional slip \citep{Cunningham2005a}\citep{Cunningham2010}. The intersection of these two fault systems provides a natural laboratory for studying strike-slip kinematics in an area with a long, varied tectonic history. Furthermore, an understanding of this critical zone will help elucidate the evolution of regional tectonic patterns.

	The 150 km wide Shargyn Basin in western Mongolia lies at the intersection of the Altai fault system and the Gobi-Altai fault system. Here, I describe the Shargyn Basin as the result of transpressional strike-slip kinematics. The kinematics consist of a compressional strike-slip stepover, a conjugate strike-slip intersection, and termination thrust splays. I also discuss the active faulting in the context of pre-existing foliation and structural weaknesses.

### 2013-05-09 writing/paper/structure.tex
almost done

\subsection{Field Mapping}
During the summer of 2012, Claire Bucholz, myself, and Esoo Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region. We spent part of the time focusing on the active faults. We performed 10-20km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}). We mapped at a broad scale, but made careful note of features indicating recent brittle deformation. Topography, fracturing, fault exposures and shifts in foliations or lithology were all important in locating faults.  Mapping was performed on paper with the assistance of satellite imagery and a GPS device. Later, the mapping was transferred to digital form.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 7in, height = 9.42in]{khantaishir_joint.png}
  \caption{A northeast-southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts paired with a Landsat image. Important locations and lithology are noted on the satellite image.}
\end{figure}

The eastern edge of the basin is uplifted with motion concentrated on two faults which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT). The cross-section across these two faults is marked on the map as A-A'. Other minor faults are identifiable on the satellite imagery. We did not investigate the minor faults in the field. Both the UKT and LKT strike approximately NNW-SSE. The UKT is the northern fault. On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence with a thick conglomerate layer followed by shales, sandstones and carbonate reef deposits. At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present. No outcrop is found on the southwest side of the UKT. At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that we interpret as the UKT itself. This zone is shown in Figure \ref{faultzones}A. Nearing this thrust trace, there is a significant increase in fracturing with strikes between 280$\circ$ and 320$\circ$. Fracture dip was similar to the dip of the UKT itself. Rotated sigmoidal tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion. S-C fabric within the deformation zone also indicated thrust motion.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 6in]{faultscollage.png}
  \caption{Photographs of exposed faults. Field workers are used as scale bars. A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section E-E'. B) An outcrop of the exposed thrust in cross-section G-G'}
  \label{faultzones}
\end{figure}

@@

Motion on the LKT may have begun after the motion on the UKT. In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons. Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed. The uplift was rapid enough that the drainage pattern was "locked in". Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the time-period before the LKT was active. The deeply drainages between the LKT and the UKT also indicate that the LKT is a later uplift. A major drainage cuts through the uplift adjacent to the LKT. Such a drainage would be unlikely to form unless it was present before the activation of the LKT, providing further evidence that uplift began on the UKT before beginning on the LKT.

@@

The Butara Thrust is exposed 25km west of cross-section B-B'. Data from a transect performed at this location is shown in cross-section X1-X1' in Appendix A. Based on the outcrop, the fault strikes E-W and dips 15$\circ$. The foot wall consists of low grade metamorphosed mudstones and sandstone. The hanging wall consists of greenschist. Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge. Kinematic indicators suggest thrust sense of motion. A photograph is shown in Figure \ref{faultzones}B. The topography rises sharply directly south of the location of the fault.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 7in, height = 8.88in]{intersection_wedge_joint.png}
  \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin. Mapped active faults are shown by solid red lines. Slip sense is shown with thrust teeth or lateral slip arrows. Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge. Cross sections C-C' and D-D' through the transtensional basin and the intersection wedge are shown. Cross-sections X2-X2' and X3-X3' are shown in Appendix A.}
  \label{intersection_wedge_joint}
\end{figure}

At the eastern corner of the intersection wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking. At the location of this bend, there is a small (2km wide) transtensional basin, shown in cross-section C-C'. This has been previously identified on satellite imagery by \citet{Cunningham2010}. We examined the area in the field. Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise. North of the basin, we found a large granite intrusion. South of the basin, there are south dipping high grade metamorphic rocks. Because of the topographic depression and local deposition, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.

@@

At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section X2-X2' in Appendix A. After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south. However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault. There is a noticeable topographic rise bending away form the main Bogd Fault. The small depositional region and the topographic rise would be unlikely without recent motion. In addition, there was a higher local degree of damage to the bedrock.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 2.0in]{shargyn_fault_streams.png}
  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section C-C'.}
  \label{shargynoffsetstreams}
\end{figure}

>>>>

\subsection{Field Mapping}
During the summer of 2012, Claire Bucholz, myself, and Yeson Erdene, a student at Mongolian University of Science and Technology performed two months of fieldwork in the Shargyn Basin region. We spent part of the time focusing on the active faults. We performed 10-20 km transects through areas with important faults, as previously identified by satellite image (mostly similar to the faults identified by \citep{Walker2007}). We mapped at a broad scale, but made careful note of features indicating recent brittle deformation. Topography, fracturing, fault exposures and shifts in foliations or lithology were all important in locating faults.  Mapping was performed on paper with the assistance of satellite imagery and a GPS device. Later, the mapping was transferred to digital form.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 8.07in]{khantaishir_joint.png}
  \caption{A northeast-southwest cross-section through the Khantaishir Range and the Khantaishir Thrusts paired with a Landsat image. Important locations and lithology are noted on the satellite image.}
\end{figure}

The eastern edge of the basin is uplifted with motion concentrated on two faults which I call the Upper Khantaishir Thrust (UKT) and the Lower Khantaishir Thrust (LKT). The cross-section across these two faults is marked on the map as A-A'. Other minor faults are identifiable on the satellite imagery. We did not investigate the minor faults in the field. Both the UKT and LKT strike approximately NNW-SSE. The UKT is the northern fault. On the northeast side of the UKT, we observed a thick fractured layer of volcaniclastics underneath a sedimentary sequence with a thick conglomerate layer followed by shales, sandstones and carbonate reef deposits. At higher elevations in the mountain range, mafic and ultramafic ophiolite-related rocks are present. No outcrop is found on the southwest side of the UKT. At the alluvium-outcrop transition, we observed a 3-5m wide deformation zone that we interpret as the UKT itself. This zone is shown in Figure \ref{faultzones}A. Nearing this thrust trace, there is a significant increase in fracturing with strikes between 280$^{\circ}$ and 320$^{\circ}$. Fracture dip was similar to the dip of the UKT itself. Rotated sigmoidal tension gashes were observed immediately above the deformation zone indicating a thrust sense of motion. S-C fabric within the deformation zone also indicated thrust motion.

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 4.53in]{faultscollage.png}
  \caption{Photographs of exposed faults. Field workers are used as scale bars. A) An outcrop of the exposed Upper Khantaishir Thrust from cross-section A-A'. B) An outcrop of the exposed thrust in cross-section X1-X1'}
  \label{faultzones}
\end{figure}

@@

Motion on the LKT may have begun after the motion on the UKT. In the small uplifted region near the trace of the LKT, there are deeply incised, but highly sinuous incised drainage canyons. Such incised drainages indicate that, before uplift began, a pre-existing sinuous drainage pattern existed. The uplift was rapid enough that the drainage pattern was "locked in". Because these drainages are aligned with the current southwestwards drainage patterns from the Khantaishir Range and the UKT, it fits with a southwestwards drainage pattern for the time-period before the LKT was active. The deep drainages between the LKT and the UKT also indicate that the LKT is a later uplift. A major drainage cuts through the uplift adjacent to the LKT. Such a drainage would be unlikely to form unless it was present before the activation of the LKT, providing further evidence that uplift began on the UKT before beginning on the LKT.

@@

The Butara Thrust is exposed 25 km west of cross-section B-B'. Data from a transect performed at this location is shown in cross-section X1-X1' in Appendix A. Based on the outcrop, the fault strikes E-W and dips 15$^{\circ}$. The foot wall consists of low grade metamorphosed mudstones and sandstone. The hanging wall consists of greenschist. Inwards, there are multiple layers of brecciated footwall and hanging wall rock and multiple layers of fault gouge. Kinematic indicators suggest thrust sense of motion. A photograph is shown in Figure \ref{faultzones}B. The topography rises sharply directly south of the location of the fault.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 6in, height = 7.68in]{intersection_wedge_joint.png}
  \caption{Landsat imagery showing the intersection wedge area at the southwestern corner of the Shargyn Basin. Mapped active faults are shown by solid red lines. Slip sense is shown with thrust teeth or lateral slip arrows. The Main Mongolian Lineament is marked by the dark black line -- dashed red and black where it overlaps with a fault. Note the two conjugate strike-slip faults, the stacked thrusts at the basin margin and the two small transtensional basins located at the eastern and northern corners of the wedge. Cross sections C-C' and D-D' through the transtensional basin and the intersection wedge are shown. Cross-sections X2-X2' and X3-X3' are shown in Appendix A.}
  \label{intersection_wedge_joint}
\end{figure}

At the eastern corner of the intersection wedge, there is a bend in the Gobi-Altai Fault from E-W striking to ENE-WSW striking. At the location of this bend, there is a small (2 km wide) transtensional basin, shown in cross-section C-C'. This has been previously identified on satellite imagery by \citet{Cunningham2010}. We examined the area in the field. Though we were not able to directly observe any deformation structures, the southern margin of this basin had a very steep topographic rise. North of the basin, we found a large granite intrusion. South of the basin, there are south dipping high grade metamorphic rocks. Because of the topographic depression and local deposition, the small basin is probably bounded by a normal fault on one side and the left-lateral Bogd Fault on the other.

@@

At the far southern corner of the thrust wedge, we examined the terminus of the Bogd Fault, shown in cross-section X2-X2' in Appendix A. After the various thrust splays further east and north, it is unlikely that there is significant relative motion so far south. However, what remains is directed into a thrust fault that strikes approximately 45 degrees off the main Bogd Fault. There is a noticeable topographic rise bending away form the main Bogd Fault. The small depositional region and the topographic rise would be unlikely without recent motion. In addition, there was a high degree of fracturing.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width = 4in, height = 2.0in]{shargyn_fault_streams.png}
  \caption{Offset stream beds along the ENE-WSW striking Shargyn Fault observed in Landsat imagery west of cross-section F-F'.}
  \label{shargynoffsetstreams}
\end{figure}
